Non-word repetition in Yélî Dnye

### 2020-09-09 manuscript.tex
update ms

\subsubsection{The Yélî language}\label{the-yuxe9luxee-language}

Yélî Dnye is an isolate language (presumed Papuan) spoken by
approximately 7000 people residing on Rossel Island, which is found at
the far end of the Louisiade Archipelago in Milne Bay Province, Papua
New Guinea. Most speakers of the language grow up speaking it
monolingually within their swidden horticulturalist communities until
they begin attending school around the age of 7; school instruction is
in English. While monolingual Yélî Dnye upbringing is common,
multilingual families are not unusual, particularly in the region around
the Catholic Mission (1953--present) where there appears to be a higher
incidence of married-in mothers from other islands ({\textbf{???}}), the
same region in which the current data were collected. Children in these
multilingual families grow up speaking Yélî Dnye and English, Tok Pisin,
and/or some other language(s) from the region.

The Yélî sound system, much like its baroque grammatical system in
general ({\textbf{???}}), is unlike any other in the region, with 90
contrastive segments, including at least two contrasts not yet been
documented elsewhere (labial-coronal double-articulations with dental
vs.~post-alveolar coronal placement in both oral and nasal stops)
({\textbf{???}}). With only four primary places of articulation
(bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts,
the phonological inventory is remarkably packed with acoustically
similar segments. The core oral stop set includes both singleton (/p/,
/t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments,
with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and
with a substantial portion of these able to be pre-nasalized or nasally
released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/,
/ṭpṇm/, /kpŋm/). Finally, a number of these can also be labialized or
palatalized on the release, or both (/pʲ/, /tʃ/, /tpʲ/, /ṭʲ/, /ṭpʲ/,
/kʲ/, /kpʲ/, /pʲʷ/, /mbʲ/, /ndʒ/, /nmdbʲ/, /ṇḍʲ/, /ṇmḍbʲ/, /mbʷ/, /ŋɡʷ/,
/mbʲʷ/, /ṭṇʲ/, /ṭpṇmʲ/, /kŋʷ/, /mʲ/, /nʲ/, /nmʲ/, /ṇʲ/, /ṇmʲ/, /mʷ/,
/ŋʷ/, mʲʷ/). Among the remaining consonants are small group of oral
continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye
may be oral or nasal, short or long. The 10 oral vowel qualities (/i/,
/ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or
long vowels, with 7 of these able to appear as short and long nasal
vowels as well (/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/). In total, Levinson
({\textbf{???}}) then counts 90 distinctive segments in Yélî Dnye (93
when including consonants that are extremely rarely used).

To do: - word shapes - typical word length

@@

@@

\caption{\label{tab:tab2-stims}NWR stimuli in orthographic and phonological representations.}
\centering
\begin{tabular}[t]{llllllllll}
\toprule
\multicolumn{2}{c}{Practice} & \multicolumn{2}{c}{Monosyll} & \multicolumn{2}{c}{Bisyll} & \multicolumn{2}{c}{Trisyll} & \multicolumn{2}{c}{Tetrasyll} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon.\\
\midrule
nopimade & nɔpimæṭɛ & dp:a & ṭp̃æ & kamo & kæmɔ & dimope & ṭimɔpɛ & diponate & ṭipɔnætɛ\\
poni & pɔni & dpa & ṭpæ & kani & kæni & diyeto & ṭijɛtɔ & nomiwake & nɔmiwækɛ\\
wî & wɯ & dpâ & ṭpɑ & kipo & kipɔ & meyadi & mɛjæṭi & todiwuma & tɔṭiwumæ\\
 &  & dpê & ṭpə & noki & nɔki & mituye & mitujɛ & wadikeno & wæṭikɛnɔ\\
 &  & dpéé & ṭpeː & nomi & nɔmi & nademo & næṭɛmɔ &  & \\
\addlinespace
 &  & dpi & ṭpi & piwa & piwæ & nayeki & næjɛki &  & \\
 &  & dpu & ṭpu & towi & tɔwi & nuyedi & nujɛṭi &  & \\
 &  & gh:ââ & ɣ̃ɑː & tupa & tupæ & pedumi & pɛṭumi &  & \\
 &  & ghuu & ɣuː &  &  & tiwune & tiwunɛ &  & \\
 &  & kpu & kpu &  &  & tumowe & tumɔwɛ &  & \\
\addlinespace
 &  & lv:ê & lβʲ̃ə &  &  & widone & wiṭɔnɛ &  & \\
 &  & lva & lβʲæ &  &  & wumipo & wumipɔ &  & \\
 &  & lvi & lβʲi &  &  &  &  &  & \\
 &  & tp:a & tp̃æ &  &  &  &  &  & \\
 &  & tpâ & tpɑ &  &  &  &  &  & \\
\addlinespace
 &  & tpê & tpə &  &  &  &  &  & \\
\bottomrule
\end{tabular}
\end{table}

@@

@@

>>>>

\subsubsection{Yélî Dnye phonological
overview}\label{yuxe9luxee-dnye-phonological-overview}

Yélî Dnye is an isolate language (presumed Papuan) spoken by
approximately 7000 people residing on Rossel Island, which is found at
the far end of the Louisiade Archipelago in Milne Bay Province, Papua
New Guinea. The Yélî sound system, much like its baroque grammatical
system (Levinson, in preparation), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar,
post-alveolar, and velar) and no voicing contrasts, the phonological
inventory is remarkably packed with acoustically similar segments. The
core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and
doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal
equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a
substantial portion of these able to be contrastively pre-nasalized or
nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/,
/kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can
further be contrastively labialized or palatalized on release, or both
(e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (in preparation)
for details). The consonantal inventory also includes a number of
non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in
Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel
qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/,
/o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or long vowels,
with 7 of these able to appear as short and long nasal vowels as well
(/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/).

In total, Yélî Dnye uses 90 distinctive segments (93 when including
consonants that are extremely rarely used), far outstripping the
phonemic inventory size of other documented Papuan languages (Levinson
(in preparation); Foley 1986) and includes at least two contrasts not
yet been documented elsewhere {[}labial-coronal double-articulations
with dental vs.~post-alveolar coronal placement in both oral and nasal
stops, Maddieson \& Levinson 2001; Ladefoged and Maddieson 1996{]}.

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with
monosyllabic words (\textasciitilde{}40\%) appearing most commonly after
that, and with tri-and-above syllablic words appearing least frequently
(\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the
current Yélî Dnye dictionary, Levinson, in preparation). The vast
majority of syllables use a CV format. A small portion of the lexicon
features words with a final closed syllable, limited to codas of -/m/,
-/p/, or -/j/ (e.g., `ndap' (Spondylus shell) /ṇṭæp/). However, in
spontaneous speech an epenthetic /ɯ/ is often inserted after word-final
coda, yielding a CVCV structure instead (e.g., `ndapî' /`ṇṭæ.pɯ/). This
process is used frequently with English loan words that have a coda
(e.g., 'ponî' (phone) /`pɔ.nɯ/). An even smaller portion of the lexicon
features words starting with a vowel (e.g., 'ala' (here) /æ.'læ/), but
these are limited to /æ/-. Finally, the lexicon features a handful of
single-vowel grammatical morphemes (see Levinson (in preparation) for
details).

While all consonants can appear in the first syllable of the word, in
bisyllabic roots, the second syllable onset consonant is almost
exclusively drawn from a subset of the simplex consonants (in order of
descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/,
/t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and
long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%),
with the first syllable of bisyllabic stems following a similar pattern.
Most vowels in the second syllable of bisyllabic stems are also short
and oral, with a potential preference for rounding harmony between the
first and second vowels. Stress typically falls on the first syllable of
bisyllabic words, on the first and third syllables of tetrasyllabic
words, and on the first or second syllable of trisyllabic words,
depending on its vocalic features (see Levinson (in preparation) for
details).

@@

Most speakers of the language grow up speaking it monolingually within
their swidden horticulturalist communities until they begin attending
school around the age of 7; school instruction is in English. While
monolingual Yélî Dnye upbringing is common, multilingual families are
not unusual, particularly in the region around the Catholic Mission
(1953--present) where there appears to be a higher incidence of
married-in mothers from other islands (Brown \& Casillas, in press), the
same region in which the current data were collected. Children in these
multilingual families grow up speaking Yélî Dnye and English, Tok Pisin,
and/or some other language(s) from the region.

@@

\caption{\label{tab:tab2-stims}NWR stimuli in orthographic and phonological representations.}
\centering
\begin{tabular}[t]{llllllllll}
\toprule
\multicolumn{2}{c}{Practice} & \multicolumn{2}{c}{Monosyll} & \multicolumn{2}{c}{Bisyll} & \multicolumn{2}{c}{Trisyll} & \multicolumn{2}{c}{Tetrasyll} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon.\\
\midrule
nopimade & nɔpimæṭɛ & dp:a & ṭp̃æ & kamo & kæmɔ & dimope & ṭimɔpɛ & dipońate & ṭipɔnætɛ\\
poni & pɔni & dpa & ṭpæ & kańi & kæni & diyeto & ṭijɛtɔ & ńomiwake & nɔmiwækɛ\\
wî & wɯ & dpâ & ṭpɑ & kipo & kipɔ & meyadi & mɛjæṭi & todiwuma & tɔṭiwumæ\\
 &  & dpê & ṭpə & ńoki & nɔki & mituye & mitujɛ & wadikeńo & wæṭikɛnɔ\\
 &  & dpéé & ṭpeː & ńomi & nɔmi & ńademo & næṭɛmɔ &  & \\
\addlinespace
 &  & dpi & ṭpi & piwa & piwæ & ńayeki & næjɛki &  & \\
 &  & dpu & ṭpu & towi & tɔwi & ńuyedi & nujɛṭi &  & \\
 &  & gh:ââ & ɣ̃ɑː & tupa & tupæ & pedumi & pɛṭumi &  & \\
 &  & ghuu & ɣuː &  &  & tiwuńe & tiwunɛ &  & \\
 &  & kpu & kpu &  &  & tumowe & tumɔwɛ &  & \\
\addlinespace
 &  & lv:ê & lβʲ̃ə &  &  & widońe & wiṭɔnɛ &  & \\
 &  & lva & lβʲæ &  &  & wumipo & wumipɔ &  & \\
 &  & lvi & lβʲi &  &  &  &  &  & \\
 &  & tp:a & tp̃æ &  &  &  &  &  & \\
 &  & tpâ & tpɑ &  &  &  &  &  & \\
\addlinespace
 &  & tpê & tpə &  &  &  &  &  & \\
\bottomrule
\end{tabular}
\end{table}

@@

\hypertarget{ref-brownIPchildrearing}{}
Brown, P., \& Casillas, M. (in press). Childrearing through social
interaction on Rossel Island, PNG. In A. J. Fentiman \& M. Goody (Eds.),
\emph{Esther Goody revisited: Exploring the legacy of an original
inter-disciplinarian} (pp. XX--XX). New York, NY: Berghahn.

@@

\hypertarget{ref-levinsoninprepydgrammar}{}
Levinson, S. C. (in preparation). \emph{A grammar of yélî dnye, the
papuan language of rossel island}. Cambridge, UK: Cambridge University
Press.
